\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2013-12-12 doc/paper.tex
minor changes to instructions for nmssm in paper.tex.
typo fix on last line in paper.tex.
minor edit to remove duplication in paper.tex.
edits to paper.tex.

\subsection{Lattice method}
If ready.
\subsection{multiple effective models at different scales}
\section{Implementing new models}
\section{Adapting the code at the C++ level}
\section{Using individual modules in your own code}
\appendix
\section{Examples}
\subsection{NMSSM}
Detailed steps to create NMSSM model file.
\begin{enumerate}
\item Create new directory \code{mkdir myNMSSM}
\item Move into this new directory \code{cd myNMSSM}
\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
\item Edit myNMSSM.m as follows
\begin{enumerate}
\item  Change Model name, author and date
\item Add singlet superfield to SuperFields
\begin{lstlisting}
SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
\end{lstlisting}
\item Modify superpotential
\begin{lstlisting}
SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} 
  DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item  Add the singlino to the list of gauge eigenstate dirac fermions
\begin{lstlisting}
  DEFINITION[GaugeES][DiracSpinors]={
  Bino ->{fB, conj[fB]},
  Wino -> {fWB, conj[fWB]},
  Glu -> {fG, conj[fG]},
  H0 -> {FHd0, conj[FHu0]},
  HC -> {FHdm, conj[FHup]},
  Fd1 -> {FdL, 0},
  Fd2 -> {0, FdR},
  Fu1 -> {FuL, 0},
  Fu2 -> {0, FuR},
  Fe1 -> {FeL, 0},
  Fe2 -> {0, FeR},
  Fv -> {FvL,0}
  FS -> {FsR,conj[FsR]}   
};
\end{lstlisting}
\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.  Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.

\begin{lstlisting}
 DEFINITION[EWSB][MatterSector]= 
{    {{SdL, SdR}, {Sd, ZD}},
     {{SvL}, {Sv, ZV}},
     {{SuL, SuR}, {Su, ZU}},
     {{SeL, SeR}, {Se, ZE}},
     {{phid, phiu, phiS}, {hh, ZH}},       <----------------
     {{sigmad, sigmau, sigmaS}, {Ah, ZA}}, <----------------
     {{SHdm,conj[SHup]},{Hpm,ZP}},
     {{fB, fW0, FHd0, FHu0, FsR}, {L0, ZN}}, <---------------
     {{{fWm, FHdm}, {fWp, FHup}}, {{Lm,UM}, {Lp,UP}}},
     {{{FeL},{conj[FeR]}},{{FEL,ZEL},{FER,ZER}}},
     {{{FdL},{conj[FdR]}},{{FDL,ZDL},{FDR,ZDR}}},
     {{{FuL},{conj[FuR]}},{{FUL,ZUL},{FUR,ZUR}}}       };

\end{lstlisting}
\end{enumerate}
\item Now we copy over particles.m \code{cp /path/to/SARAH/Models/MSSM/particles.m  ./}
\item Edit particles.m as follows
\begin{enumerate}  
\item First we add to particleDefinitions[GaugeES]:
\begin{lstlisting}
 ParticleDefinitions[GaugeES] = {
      {SdL,  { Description -> "Left Down-Squarks"}},
      {SdR,  { Description -> "Right Down-Squarks"}},
      ...........
      ...........
      ...........
      {H0,   { Description -> "Neutral Higgsinos"}},
      {HC,   { Description -> "Charged Higgsinos"}},
      {SsR, { Description -> "Singlet"}},         <---------------
      {FS,   { Description -> "Singlino" }},      <---------------
      {Fd1,  { Description -> "Dirac Left Down-Quark"}},
      {Fd2,  { Description -> "Dirac Right Down-Quark"}},
      ............
      ............
      };
\end{lstlisting}
Note that we are brief in information here becasue there is already some information contained in the default particles.m file in the model directory:
\begin{lstlisting} 
{{ Description -> "Singlet", 
                 PDG -> {0},
                 PDG.IX -> 0,
                 Width -> 0, 
                 Mass -> Automatic,
                 FeynArtsNr -> 98,
                 LaTeX -> "S",
                 OutputName -> "s"

@@

{{ Description -> "Singlet-VEV", 
			 LaTeX -> "v_s",
             Value -> None, 
             LesHouches -> {NMSSMRUN,5},
             OutputName-> vS         }},
\end{lstlisting}
\item Navigate to Flexible SUSY directory  \code{cd /path/to/FlexibleSUSY/}
\item Create Z3NMSSM directory \code{./createmodel --models=Z3NMSSM:myNMSSM}
\item Configure with the new model \code{./configure --with-models=Z3NMSSM}
\item The default FlexibleSUSY model file for the MSSM will now be in the models/NMSSM2 directory.  Edit this for the NMSSM as follows:
  \begin{enumerate} 
    \item Add new input parameter $\lambda$.
      \begin{lstlisting}
        EXTPAR = { {61, LambdaInput} };
      \end{lstlisting}
     \item Also add this parameter to the DefaultParameterPoint.
      \begin{lstlisting}
       DefaultParameterPoint = {
    {m0, 125},
    {m12, 500},
    {TanBeta, 10},
    {Sign[\[Mu]], 1},
    {Azero, 0}
    {LambdaInput, 0.1}
    };
  \end{lstlisting}
    \item Name the parameters you wish to be output by the EWSB conditions.  Note you should use the names given for them in parameters.m.    A common choice in the NMSSM is:
\begin{lstlisting}
EWSBOutputParameters = {vS, \[Kappa], ms2};
\end{lstlisting}
which will work for the semi-constrained NMSSM.
\item choose the Highscale and Highscale first guess, if the default is different from what you wish to use.  
\begin{lstlisting} 
 HighScale = g1 == g2;
 HighScaleFirstGuess = 1.0 10^16;
\end{lstlisting}
\item Add how the parameters found in parameters.m that are to be set at the highscale:
\begin{lstlisting} 
HighScaleInput={
   {T[Ye], Azero*Ye},
   {T[Yd], Azero*Yd},
   {T[Yu], Azero*Yu},
   {mHd2, m0^2},
   {mHu2, m0^2},
   {mq2, UNITMATRIX[3] m0^2},
   {ml2, UNITMATRIX[3] m0^2},
   {md2, UNITMATRIX[3] m0^2},
   {mu2, UNITMATRIX[3] m0^2},
   {me2, UNITMATRIX[3] m0^2},
   {MassB, m12},
   {MassWB,m12},
   {MassG,m12},
{\[Lambda], LambdaInput},
   {T[\[Kappa]], Azero \[Kappa]},
   {T[\[Lambda]], Azero LambdaInput}
};
\end{lstlisting}
\item set lowscale boundary condititions if you want something different from the default.  The defaults are:
\begin{lstlisting} 
 LowScale = SM[MZ];
 LowScaleFirstGuess = SM[MZ];
\end{lstlisting}
\item Change the LowscaleInput if necessary.  In this example we keep the same choice as in the MSSM
  \begin{lstlisting} 
 LowScaleInput={
   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
};
  \end{lstlisting}

@@

\end{thebibliography}
\end{document
}

>>>>

\subsection{Lattice method}
If ready.
\subsection{multiple effective models at different scales}
\section{Implementing new models}
\section{Adapting the code at the C++ level}
\section{Using individual modules in your own code}
\appendix
\section{Examples}
\subsection{NMSSM}
Detailed steps to create NMSSM model file.
\begin{enumerate}
\item Create new directory \code{mkdir myNMSSM}
\item Move into this new directory \code{cd myNMSSM}
\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
\item Edit myNMSSM.m as follows
\begin{enumerate}
\item  Change Model name, author and date
\item Add singlet superfield to SuperFields
\begin{lstlisting}
SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
\end{lstlisting}
\item Modify superpotential
\begin{lstlisting}
SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} 
  DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item  Add the singlino to the list of gauge eigenstate dirac fermions
\begin{lstlisting}
  DEFINITION[GaugeES][DiracSpinors]={
 .............
  FS -> {FsR,conj[FsR]}   
};
\end{lstlisting}
\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.  Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.

\begin{lstlisting}
 DEFINITION[EWSB][MatterSector]= 
{   ................
    ................
     {{phid, phiu, phiS}, {hh, ZH}},      
     {{sigmad, sigmau, sigmaS}, {Ah, ZA}},
    ................
     {{fB, fW0, FHd0, FHu0, FsR}, {L0, ZN}},
     ...............
};

\end{lstlisting}
\end{enumerate}
\item Now we copy over particles.m \code{cp /path/to/SARAH/Models/MSSM/particles.m  ./}
\item Edit particles.m as follows
\begin{enumerate}  
\item First we add to particleDefinitions[GaugeES]:
\begin{lstlisting}
 ParticleDefinitions[GaugeES] = {
      {SsR, { Description -> "Singlet"}},        
      {FS,   { Description -> "Singlino" }},    
      ............
      };
\end{lstlisting}
Note that we are brief in information here becasue there is already some information contained in the default particles.m file in the model directory:
\begin{lstlisting} 
{{ Description -> "Singlet", 
                 PDG -> {0},
                 PDG.IX -> 0,
                 Width -> 0, 
                 Mass -> Automatic,
                 FeynArtsNr -> 98,
                 LaTeX -> "S",
                 OutputName -> "s"

@@

{{ Description -> "Singlet-VEV", 
			 LaTeX -> "v_s",
             Value -> None, 
             LesHouches -> {NMSSMRUN,5},
             OutputName-> vS         }},
\end{lstlisting}
\item Navigate to Flexible SUSY directory  \code{cd /path/to/FlexibleSUSY/}
\item Create Z3NMSSM directory \code{./createmodel --models=Z3NMSSM:myNMSSM}
\item Configure with the new model \code{./configure --with-models=Z3NMSSM}
\item The default FlexibleSUSY model file for the MSSM will now be in the models/NMSSM2 directory.  Edit this for the NMSSM as follows:
  \begin{enumerate} 
    \item Add new input parameter $\lambda$ to EXTPAR and remove the sign of $\mu$ from MINPAR.
      \begin{lstlisting}    
        MINPAR = { {1, m0},
           {2, m12},
           {3, TanBeta},
           {5, Azero} };

  EXTPAR = { {61, LambdaInput} };
      \end{lstlisting}
     \item Also add this parameter to the DefaultParameterPoint.
      \begin{lstlisting}
       DefaultParameterPoint = {
    {m0, 200},
    {m12, 200},
    {TanBeta, 10},
    {Azero, -500}
    {LambdaInput, 0.1}
    };
  \end{lstlisting}
    \item Name the parameters you wish to be output by the EWSB conditions.  Note you should use the names given for them in parameters.m.    A common choice in the NMSSM is:
\begin{lstlisting}
EWSBOutputParameters = {vS, \[Kappa], ms2};
\end{lstlisting}
which will work for the semi-constrained NMSSM.
\item choose the Highscale and Highscale first guess, if the default is different from what you wish to use.  
\begin{lstlisting} 
 HighScale = g1 == g2;
 HighScaleFirstGuess = 1.0 10^16;
\end{lstlisting}
\item Add how the parameters found in parameters.m that are to be set at the highscale:
\begin{lstlisting} 
HighScaleInput={
   {T[Ye], Azero*Ye},
   {T[Yd], Azero*Yd},
   {T[Yu], Azero*Yu},
   {mHd2, m0^2},
   {mHu2, m0^2},
   {mq2, UNITMATRIX[3] m0^2},
   {ml2, UNITMATRIX[3] m0^2},
   {md2, UNITMATRIX[3] m0^2},
   {mu2, UNITMATRIX[3] m0^2},
   {me2, UNITMATRIX[3] m0^2},
   {MassB, m12},
   {MassWB,m12},
   {MassG,m12},
{\[Lambda], LambdaInput},
   {T[\[Kappa]], Azero \[Kappa]},
   {T[\[Lambda]], Azero LambdaInput}
};
\end{lstlisting}
\item set lowscale boundary condititions if you want something different from the default.  The defaults are:
\begin{lstlisting} 
 LowScale = SM[MZ];
 LowScaleFirstGuess = SM[MZ];
\end{lstlisting}
\item Change the LowscaleInput if necessary.  In this example we keep the same choice as in the MSSM
  \begin{lstlisting} 
 LowScaleInput={
   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
};
  \end{lstlisting}

@@

\end{thebibliography}
\end{document}
